---
title: "panelappR"
author: "[David Montaner](http://www.dmontaner.com)"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: no
    fig_width: 5
    fig_height: 5
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

`panelappR` is an R library to download
[PanelApp](https://panelapp.extge.co.uk)
data.




Install
================================================================================

You can install the library doing:

    library (devtools)
    install_github ("dmontaner/panelappR/pkg")




Usage
================================================================================

Load the package

```{r}
library (panelappR)
```


Download the data

```{r, results = 'hide'}
pad <- paDownload ()
```

`pad` is a list containing the following data frames:

- panels
- disorders
- genes
- evidences
- publications
- phenotypes
- ensembl


```{r}
lapply (pad, head)
```


Annotate the genes using the latest Ensembl information.

```{r, results = 'hide', warning = FALSE}
pad <- paAnnotate (pad)
```

The the Ensembl information is now incorporated  in new columns of the data.frame `pad$ensembl`

```{r}
head (pad$ensembl)
```

Use `append = FALSE` if you want to get the information in an independent data.frame.
Use `latestVersion = FALSE` if you want to annotate using Ensembl 80,
the version in which PanelApp was firstly developed.


Export into an `xlsx` file:

```{r}
paExport.xlsx (pad, file = "panelap_data.xlsx")
```

