---
title: "panelappR"
author: "[David Montaner](http://www.dmontaner.com)"
date: "Created: 2017-01-18 | Revised: 2017-03-13 | Compiled: `r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: no
    fig_width: 5
    fig_height: 5
vignette: >
  %\VignetteIndexEntry{panelappR intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

--------------------------------------------------------------------------------

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options (width = 80)
```

\  


`panelappR` is an R library to download
[PanelApp](https://panelapp.extge.co.uk)
data.


Install
================================================================================

You can install the library doing:

    install.packages ("devtools")
    library (devtools)
    setRepositories (ind=1:2)
    install_github ("dmontaner/panelappR/pkg")


Usage
================================================================================

Load the package

```{r}
library(panelappR)
```

Download the data

```{r, results = 'hide'}
pad <- paDownload()
```

`pad` is a list containing the following data frames:

- panels
- disorders
- genes
- evidences
- publications
- phenotypes
- ensembl


```{r}
names(pad)
head(pad$genes)
```

Annotate the genes using the latest Ensembl information.

```{r, results = 'hide', warning = FALSE}
pad <- paAnnotate(pad)
```

The the Ensembl information is now incorporated  in new columns of the data.frame `pad$ensembl`

```{r}
head(pad$ensembl)
```

Use `append = FALSE` if you want to get the information in an independent data.frame.
Use `latestVersion = FALSE` if you want to annotate using Ensembl 80,
the version in which PanelApp was firstly developed.


Export into an `xlsx` file:

```{r}
paExport.xlsx(pad, file = "panelap_data.xlsx")
```

Export into an `SQLite` file:

```{r, results = 'hide'}
paExport.sqlite(pad, file = "panelap_data.sqlite")
```

```{r, echo = FALSE, results = 'hide'}
unlink("panelap_data.xlsx")
unlink("panelap_data.sqlite")
```
